---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/poissonconsulting/teckfish/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/poissonconsulting/teckfish/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/teckfish/branch/main/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/teckfish?branch=main)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

# teckfish


## Introduction

`teckfish` is an R package for Teck Fish and Aquatic Sciences Projects.
It is intended to be used by Teck staff and contractors and anyone else
who finds it useful.

## Installation

```r
# install.packages("devtools")
devtools::install_github("poissonconsulting/teckfish")
```

## Demonstration

### Growing Season Degree Days

`gsdd_cf()` takes a numerical vector of mean daily temperature values 
in centigrade and calculates the growing season degree days (GSDD) based 
on Coleman and Fausch's (2007) definition.

```{r}
library(teckfish)
gsdd_cf(teckfish::simulated_data$synthetic)
```
