---
title: "Classify water temperature time series values into reasonable, questionable, or erroneous"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classify water temperature time series values into reasonable, questionable, or erroneous}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(teckfish)
library(dplyr)
library(tibble)
library(ggplot2)
library(scales)
```

Raw logger temperature time series data can be run through `classify_water_temp_data()` to classify each value as reasonable, questionable, or erroneous. 
Only a single deployment can be classified at a time.
Users are responsible for ensuring that the data are in the correct format before running this function.

The values are classified following these rules:

1. If the temperature is greater then or equal to or less then or equal to the min and max parameters (`questionable_min`, `questionable_max`, `erroneous_min`, `erroneous_max`). 
2. If the rate of change (degrees per hour) between the adjacent lag value is greater then or equal to the rate parameters (`questionable_rate`, `erroneous_rate`).
3. A value adjacent to a questionable/erroneous value are coded as questionable/erroneous. 
4. A time buffer (hours) is added to the questionable/erroneous values such that adjacent values within or equal to the buffer (`questionable_buffer`, `erroneous_buffer`) are coded as questionable/erroneous.
5. Reasonable values identified between questionable or erroneous values are coded as questionable if the gap between the nearest questionable or erroneous value is less then or equal to the `gap_range` parameter.
6 If a temperature value is missing the function will ignore the value and not classify it.

## Usage 

```{r}
data <- tibble::tribble(
    ~temperature_date_time, ~water_temperature,
    "2021-05-07 00:00:00",  40.145,
    "2021-05-07 01:00:00",  20.257,
    "2021-05-07 02:00:00",  6.257,
    "2021-05-07 03:00:00",  5.245,
    "2021-05-07 04:00:00",  2.567,
    "2021-05-07 05:00:00",  2.687,
    "2021-05-07 06:00:00",  2.175,
    "2021-05-07 07:00:00",  2.817,
    "2021-05-07 08:00:00",  2.917,
    "2021-05-07 09:00:00",  2.817,
    "2021-05-07 10:00:00",  2.867,
    "2021-05-07 11:00:00",  2.817,
    "2021-05-07 12:00:00",  2.917,
    "2021-05-07 13:00:00",  2.817,
    "2021-05-07 14:00:00",  2.867,
    "2021-05-07 15:00:00",  2.877,
    "2021-05-07 16:00:00",  3.012,
    "2021-05-07 17:00:00",  3.045,
    "2021-05-07 18:00:00",  3.124,
    "2021-05-07 19:00:00",  3.268,
    "2021-05-07 20:00:00",  3.789,
    "2021-05-07 21:00:00",  4.104,
    "2021-05-07 22:00:00",  4.178
  ) |>
    dplyr::mutate(
      temperature_date_time = as.POSIXct(temperature_date_time, tz = "UTC")
    )

  classified_data <- classify_water_temp_data(data)
  
  classified_data
```

```{r, out.width = "100%", fig.width = 10, fig.height = 5}
ggplot(classified_data) +
  geom_point(
    aes(
      x = temperature_date_time,
      y = water_temperature, 
      color = status_id
    ),
    alpha = 0.5
  ) +
  labs(
    y = "Temperature",
    x = "Datetime"
  ) +
  theme_bw() +
  scale_x_datetime(
    breaks = date_breaks("4 hour"),
    labels = date_format("%H:%M")
  ) +
  scale_y_continuous(
    breaks = seq(-5, 45, 5),
    limits = c(-2, 42)
  ) +
  scale_color_manual(
    "Status ID",
    values = c(
      "reasonable" = "black",
      "questionable" = "orange",
      "erroneous" = "red"
    )
  )
```

### Changing temperature min and max parameters 


```{r}
classify_water_temp_data(
  data,
  questionable_max = 1,
  erroneous_max = 20
)
```

### Changing rate of change parameters

```{r}
classify_water_temp_data(
  data,
  questionable_rate = 1,
  erroneous_rate = 10
)
```

### Removing the time buffer

```{r}
classify_water_temp_data(
  data,
  questionable_buffer = 0,
  erroneous_buffer = 0
)
```

### Removing the gap filling

```{r}
classify_water_temp_data(
  data,
  gap_range = 0
)
```

## Process

1. Read in a logger file
2. Create date time column and name it `temperature_date_time`
3. Ensure the temperature column is named `water_temperature`
4. Confirm data is from a single deployment
5. Run data through `classify_water_temp_data()` function
6. Save data to a csv file

```{r, eval=FALSE}
# install.packages("tidyverse")
# install.packages("devtools")
# devtools::install_github("poissonconsulting/teckfish")

library(teckfish)
library(tidyverse)

# read in logger file
data <- readr::read_csv("path/to/file.csv")

# create date time column
data <- data |>
  dplyr::mutate(
    temperature_date_time = as.POSIXct(temperature_date_time, tz = "UTC")
  )

# ensure temperature column is named water_temperature
data <- data |>
  dplyr::rename(
    water_temperature = temperature
  )

# confirm data is from a single deployment
data <- 
  data |>
  dplyr::filter(
    deployment_id == "12345"
  )

# run data through classify_water_temp_data() function
classified_data <- classify_water_temp_data(data)

# save data to a csv file
readr::write_csv(classified_data, "path/to/file.csv")
```
